### 2024-08-26 msiddiqui60_HW1_report.tex
theoretical questions committed

\documentclass[twoside,10pt]{article}
\usepackage{amsmath,amsfonts,amsthm,fullpage}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}

\begin{document}

\title{ISYE 6740 HW 1}
\author{Mohammed Khalid Siddiqui}
\date{}

\section{Concept Questions}

\begin{enumerate}

\item In classical machine learning, Supervised learning includes both Classification and Regression
tasks whereas Unsupervised learning tends to be associated with Clustering. The primary difference between
Supervised and Unsupervised learning is that the former uses labeled data to model associations between
inputs and outputs whereas the latter does not use labeled data, instead opting to uncover associations
through similarity metrics like euclidian distance in the example of clustering. Two benefits or use-cases
of supervised learning can be to predict or forecast outputs like house prices based on a series of inputs
such as location, size, number of bedrooms and other input features as well as classify customers at risk
of churning based on a series of input data. The first task was a regression, and the second a classification.
Both are examples of supervised learning as they require input and output labels to function and attempt
to predict the output label based on the input. This serves myriad of use cases across industry. Drawbacks of
supervised learning include the need for output labels and model fitting to known data. Adding labels
such as in the case of an image dataset can be extremely time consuming. Additionally, the use of labels means
the analyst must have some idea of what they are modeling and trying to uncover. It can be harder to explore
data if you are limited to seeking a specific output which you must define. Unsupervised learning by contrast
carries the strength that allows one to explore similarities between data without explicitly defining outputs.
This focus on finding similar associations generally means that unsupervised learning may not require the training
samples in the same way supervised learning tasks do. One drawback of unsupervised learning is that unlike
supervised learning, because you aren't using labeled outputs, it may be harder to measure success via metrics like
accuracy or mean squared error etc. For this reason, unsupervised learning may be stronger suited to
exploratory data analysis tasks or creating associations of like groups without penalties for "bad" groupings.

\item Per the hint provided in the homework to use Hamming distance for the one-hot encoded version of the categorical
variables and the Google definition of (and lesson information on) Hamming distance being the number
of positions where the corresponding symbols are different, we can intuit that for any categorical variable
that has been one-hot eoncded, the Hamming distance equivalent value would be 0 if the values are the same,
and 2 if they are different, irrespective of the number of possible values available. The reason is because,
each categorical variable will only allow an observation to select one option. For example, let's assume
the second feature corresponds to 'building type' and let's say that there are n = 4 building types: "house",
"office", "apartment", "condo". Once we have one-hot encoded this feature, we will have 4 columns in the place of
the original. Each column will encode either a 1 or 0 depending on whether that value is present or not.
Going back to the example presented to us, assume the column order is the same as the list above. For those
four columns, x1 will be encoded as follows 1,0,0,0. x2 also has a building type of house and similarly, the
one-hot encoded features will appear as 1,0,0,0. If x3 were to be of type "office", the encoding would
be 0,1,0,0. Any observation can have only 1 building type. The Hamming distance for building type between x2
and x3 is 2 because two of the columns are different. The same logic can be applied to the 'city' feature.
Each observation can only have 1 city and the Hamming distance will be either 0 if agreed or 2 if disagreed.
For our numerical variable which appears to be price, we can use Euclidean distance which is
\begin{equation}
d(p,q) = \sqrt{\sum_{i=1}^{n} \left( p_i - q_i \right)^2}
\end{equation}
We will need to incorporate all three distances, the two hamming distances and the one euclidian distance
to create our clustering distance formula. This will require us to prioritize, standardize, and weight
our distances appropriately to arrive at a final distance in an reasonable manner. For the sake of simplicity,
for this problem, let's assume that city, building type, and cost are all equally important in calculating
similarity between buildings. To acheive this, we will standardize the Euclidian distance d(p,q) above
such that values are between 0 and 2. We use 2 as our upper bound because that is the maximum Hamming distance
acheivable for building type and city respectively.
I will use the following calculation to normalize the Euclidian function:
\begin{equation}
d_{\text{norm}}(x, y) = \frac{2 \cdot d(x, y)}{d_{\text{max}}}
\end{equation}
where d_max is the largest distance acheivable between building prices.
The final distance equation can act as an ensemble or combination of our three calculated distances
Hamming_buildingtype, Hamming_city, and d_norm. For a simple preliminary model, additive distance
should suffice:
\begin{equation}
d_total = Hamming_buildingtype + Hamming_city + d_norm
\end{equation}
Upon running an initial iteration of the clustering algorithm, it may be helpful to consider weighting the
different distance components or even using different transformations or formula.

\item For this problem we are comparing the functions inside the arg min function to prove their
equivalence, or specifically that given the same inputs, they will product the same outputs. The argmin
function in this instance will identify and return the value of j (the cluster a data point should be assigned to)
between 1 and k that minimizes the square of the Euclidean norm or L2 distance (As discussed in the lectures)
between the data point x_i and centroid c_j. Re-writing the Euclidean distance will be
\begin{equation}
\|x_i - c_j\|^2 = (c_j^\top c_j - 2c_j^\top x_i) + x_i^\top x_i

c_j^\top c_j - 2c_j^\top x_i

\frac{1}{2} c_j^\top c_j - c_j^\top x_i
\end{equation}
In the above we first expanded the squared L2 norm distance and re-wrote our features in vector format.
Realizing we're applying the arg min function with respect to change in j, we dropped the "constant" terms
which are identical across the two and redundant for any given data point. Lastly, knowing that the result
of the argmin function is irrespective of scale of the inputs: meaning the output of argmin(1,3) is the
same as the output of argmin(2,6) or argmin(10, 30), we divided by a factor of 2 to make the expression
equivalent to the resulting one in the homework problem.

\item K-means algorithm operates by setting initial centroids for each cluster, then identifying the points
which belong to those clusters, then updating those centroids, then identifying the points which belong to
the new clusters. This iterative process may not always converge to the "global optimal solution". I believe
in the lecture most k-means are identified as non-convex and may get trapped in local minima or converge to
local optima. This is interestingly also an issue sometimes with the gradient descent algorithm used to optimize
neural networks. In the k-means because there isn't an oscillating movement and instead a monotonic movement,
a converge is almost all but guaranteed but it won't necessarily be a globally optimal convergence. To this end,
it may be helpful to run k-means with multiple iterations using different initial centroids.

\item K-means is guaranteed to converge, if only to a local optimum not global, because there are a finite
number of k clusters ot iterate over and because the objective function, as discussed in the lectures, is
monotonically decreasing. As the objective function does not oscillate, it is guaranteed to converge to some
minimum. Due to the finite number of data points, the clusters can only update via the objective function
to a finite posisble number of combinations. Out of the three reasons provided: finite num k clusters, finite
number of data points, and monotonically decreasing objective functions, it is the objective function which
is the most compelling reason for convergence as an oscillating function could theoretically result in a loop
that continues infinitely.

\item Considering the graph provided in the question, we will first need to create an adjacency matrix,
then will we create the degree matrix. Following the degree matrix, we will then calculate the
graph Laplacian using the convenient formula L = Degree Matrix - Adjacency Matrix. Using the Graph
Laplacian, we will use the numpy library to then calculate the eigenvalues and their respective eigenvectors.
The number of 0 eigenvalues correspeonds to the number of connected components in the graph and looking
at the corresponding eigenvectors for those eigenvalues will show us which nodes belong to which connected
components because they will possess the same entries in the eigenvector. Please reference the ipynb file
for the code. Below, I'll add the relevant matrices and eigenvalues and eigenvectors printed. You'll see
the first three entries in the first eigenvector represent the first three nodes and belong to the same
connected component whereas the second eigenvector contains the last two nodes belong to the same connected
component.

\textbf{Adjacency Matrix}
\[
\begin{matrix}
0 & 1 & 1 & 0 & 0 \\
1 & 0 & 1 & 0 & 0 \\
1 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 1 & 1
\end{matrix}
\]

\textbf{Degree Matrix}
\[
\begin{matrix}
2 & 0 & 0 & 0 & 0 \\
0 & 2 & 0 & 0 & 0 \\
0 & 0 & 2 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 & 1
\end{matrix}
\]

\textbf{Graph Laplacian: D - A}
\[
\begin{matrix}
2 & -1 & -1 & 0 & 0 \\
-1 & 2 & -1 & 0 & 0 \\
-1 & -1 & 2 & 0 & 0 \\
0 & 0 & 0 & 1 & -1 \\
0 & 0 & 0 & -1 & 1
\end{matrix}
\]

\textbf{Eigenvalues}
\[
\begin{matrix}
0 & 0 & 2 & 3 & 3
\end{matrix}
\]

\textbf{Relevant Eigenvectors (correspond to eigenvalues = 0)}
\[
\begin{matrix}
-0.57735027 & 0  \\
-0.57735027 & 0  \\
-0.57735027 & 0  \\
0 & 0.70710678  \\
0 & 0.70710678
\end{matrix}
\]

\end{enumerate}